A Plasma Torus Around a Young Low-Mass Star

### 2025-01-26 paper/ms.tex
update

\subsection{Observations \& Data Reduction}

@@

\subsection{Spectroscopic Variability}

@@

\subsection{Stellar Parameters}

For TIC~141146667, we measured the velocities from each of our HIRES
spectra using regions near the K~I (7700\,\AA) resonance line and
three TiO bandheads (5160\,\AA, 5450\,\AA, and 5600\,\AA).  These
regions were selected because they provided the best matches between
the synthetic and observed spectra.  The resulting redshift
measurements were averaged together.  The star's 130$\pm$4\,\kms\
equatorial velocity and near-edge on viewing orientation yielded a
line broadening $\Delta \lambda$$\approx$3\,\AA.  We used the scatter
of resulting velocity measurements between orders to assign the RV
uncertainty at each epoch.  The uncertainty-weighted mean systemic
velocity over all epochs on UT 2024-02-17 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are given in Table~\ref{tab:rv}.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
This radial velocity time-series rules out any significant periodicity
in the radial velocities at the rotation period exceeding
semi-amplitudes of 2.85\,\kms\ (at 3$\sigma$ confidence).
% from K_to_msini.py, assuming Mstar = 0.18
This in turn sets an upper limit on the mass of any putative
companions at this period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found over 90\% of CPVs to be associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is one of the exceptions.  We calculated the probability
of TIC~141146667 being part of any of the nearby known groups using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  Based on that model,
it appears to be a field star at $>$99.9\% confidence.  Searching the
local vicinity of the star for neighbors with similar projected on-sky
velocity \cite{Tofflemire2021} similarly yields no strong candidates
for co-moving stars --- defined as those with projected tangential
velocities $\Delta v_{\rm T}$$<$5\,\kms --- that share its isochronal
youth.

{\it Age: Isochrones}--- $\sim$TucHor or younger.  Maybe 30-35myr?  NB
  color puts it at ~M4.5 (main seq).  BP-RP=3.3..
    % uhhh careful w what you mean by "isochrone"

{\it Age: Lithium}--- Frankly, it's not obviously there!  (but maybe
this is bc of the blending, and a pEW would show smth small)

This is weird.  Are you 100\% sure it's not a binary?  Ups the need for
the RV analysis.

{\it Binarity}

>>>>

\subsection{Observations \& Data Reduction}\phantom{+}

@@

\subsection{Stellar Parameters}\phantom{+}

For TIC~141146667, we measured the velocities from each of our HIRES
spectra using regions near the K~I (7700\,\AA) resonance line and
three TiO bandheads (5160\,\AA, 5450\,\AA, and 5600\,\AA).  These
regions were selected because they provided the best matches between
the synthetic and observed spectra.  The resulting redshift
measurements were averaged together.  The star's 130$\pm$4\,\kms\
equatorial velocity and near-edge on viewing orientation yielded a
line broadening $\Delta \lambda$$\approx$3\,\AA.  We used the scatter
of resulting velocity measurements between orders to assign the RV
uncertainty at each epoch.  The uncertainty-weighted mean systemic
velocity over all epochs on UT 2024-02-17 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are given in Table~\ref{tab:rv}.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

{\it No Evidence For Binarity}---Any significant periodicity in the
radial velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companions at the four
hour period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE), a proxy for the goodness of fit for a single-source
astrometric model to the Gaia astrometry, is 1.23, which is within the
usual range for apparently single sources.  There are no resolved
sources in the Gaia DR3 point source catalog.  Finally, we checked the
TESS light curve for evidence of secondary photometric periods by
subtracting the mean CPV signal over each sector and performing a
phase-dispersion minimization analysis
\cite{Stellingwerf1978,2021zndo...1011188B}.  There were no secondary
periods in the TESS data.  Previous work \cite{Bouma2024} has shown
that about 30\% of CPVs show evidence for excess noise above the Gaia
single-source astrometric model, and about 40\% of CPVs show evidence
for unresolved binary companions based on the presence of secondary
photometric periods.  This agrees with analyses showing that
multi-periodic low-mass stars are generally unresolved binaries
\cite{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that for field M dwarfs \cite{Winters2019},
pointing to a weak or non-existent connection between the CPV
phenomenon and binarity.  For TIC~141146667 specifically, we find 
no evidence for stellar multiplicity.

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found that over 90\% of CPVs are associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is one of the exceptions.  We calculated the probability
of TIC~141146667 being part of any of the nearby known groups using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  Based on that model,
it appears to be a field star at $>$99.9\% confidence.  We also searched the
local vicinity of TIC~141146667 for neighbors with similar projected on-sky
velocities using \texttt{comove} \cite{Tofflemire2021}.  This yielded
no strong candidates for co-moving stars --- defined as those with
projected tangential velocities $\Delta v_{\rm T}$$<$5\,\kms ---
that share its isochronal youth.

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main-sequence M dwarf, similar
to all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.
The star's proximity ($d$=58\,pc) and its high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.
{\bf Figure~X} shows the location of the star in the color--absolute
magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
Upper-Centaurus-Lupus (UCL), IC~2602, and the Pleiades.
To make this diagram, {\bf TODO}we adopted the USco and UCL members
from X, the IC~2602 members from Y, and the Pleiades members from Z,
and applied the same reddening corrections as in \cite{Bouma2022}.
To measure the isochrone age of TIC~141146667, we then followed the 
empirical approach proposed by \cite{Gagne2020} and
implemented in \cite{Bouma2022}.
The idea is to estimate the star (or star cluster's) age by
interpolating between the observed isochrones of star clusters with
ages that have been otherwise calibrated, for instance using
the lithium depletion-boundary method.
While this circumvents known modelling issues in molecular opacities
and starspot coverage for young M dwarfs,
the ages returned by this procedure depend on the ages
assumed for each reference cluster.
In this case, the most important reference is UCL, since TIC~141146667
overlaps directly on its isochrone on the CAMD.
We adopted the 16~Myr age for UCL from \cite{PecautMamajek2016}, which
is consistent with ages determined by more recent sub-clustering
analyses \cite{Ratzenbock2023} for the largest sub-populations.
This in turns yields an isochronal age for TIC~141146667 of
$t_{\rm iso}$=$16^{+19}_{-6}$\,Myr.
Qualitatively, this range of uncertainty is consistent with no stars
in the $\approx$40\,Myr IC~2602 cluster with similar redness as
TIC~141146667 being equally bright.

{\it Age: Lithium}--- Frankly, it's not obviously there!  (but maybe
this is bc of the blending, and a pEW would show smth small)
%TODO TODO TODO

{\it Age: Summary}---The main indicators for the youth of
TIC~141146667 are {\it i)} that it is a complex periodic variable;
{\it ii)} that it is 1.5 magnitudes brighter than main sequence stars
of the same color, while showing no indicators for binarity; {\it
iii)} weak lithium absorption.

\subsection{Spectroscopic Variability}\phantom{+}

@@
